---
format:
  typst:
    papersize: us-letter
    margin:
      x: 2.2cm
      y: 2.4cm
    mainfont: "Inter"
    font-paths: fonts
    fontsize: 10.5pt
params:
  data_path: "data.rds"
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: "ragg_png"
    dpi: 150
---

```{r setup}
#| include: false
d <- readRDS(params$data_path)
e <- d$est
colors <- d$colors
source("report_charts.R")

usd <- function(x) paste0("$", format(round(x), big.mark = ",", scientific = FALSE))
typst <- function(...) cat("```{=typst}\n", paste0(..., collapse = "\n"), "\n```\n", sep = "")
esc <- function(x) gsub("([#$\\[\\]@])", "\\\\\\1", x, perl = TRUE)
```

```{r cover}
#| output: asis
typst(sprintf('
#show heading: set text(fill: rgb("%s"))
#v(2.2cm)
#image("logo.svg", width: 52pt)
#v(0.8cm)
#text(size: 26pt, weight: 700, fill: rgb("%s"))[Software Development \\ Cost Estimate]
#v(0.3cm)
#text(size: 11pt, fill: rgb("%s"))[Prepared %s · %s]
#v(1.6cm)
#line(length: 100%%, stroke: 0.5pt + rgb("%s"))
#v(1.2cm)
#text(size: 12pt, fill: rgb("%s"))[Estimated development cost]
#linebreak()
#text(size: 38pt, weight: 700, fill: rgb("%s"))[%s]
#v(0.2cm)
#text(size: 11pt, fill: rgb("%s"))[Estimate range (±30%%): %s – %s]
#v(0.6cm)
#text(size: 10pt, fill: rgb("%s"))[%s · %s people · %s person-months]
#v(1fr)
#text(size: 8.5pt, fill: rgb("%s"))[Generated by Shiny Cost Estimator v%s. Model: COCOMO II.2000 with modern-framework calibration.]
#pagebreak()
',
  colors$primary_dark, colors$primary_dark, colors$ink_muted,
  format(d$generated_on, "%B %d, %Y"),
  esc(d$project_name),
  colors$border,
  colors$ink_muted, colors$primary, esc(usd(e$cost_usd)),
  colors$ink_muted, esc(usd(e$estimate_range$low)), esc(usd(e$estimate_range$high)),
  colors$ink,
  paste0(e$final_schedule_months, "-month schedule"),
  format(e$final_people, big.mark = ","),
  format(e$effort_person_months, big.mark = ","),
  colors$ink_muted, d$app_version
))
```

# Estimate Summary

```{r kpis}
#| output: asis
tco_cell <- if (!is.null(e$maintenance)) {
  sprintf('[#text(size: 8pt, fill: rgb("%s"))[TOTAL COST (TCO)] #linebreak() #text(size: 15pt, weight: 700)[%s]]',
          colors$ink_muted, esc(usd(e$maintenance$tco)))
} else {
  sprintf('[#text(size: 8pt, fill: rgb("%s"))[EFFORT] #linebreak() #text(size: 15pt, weight: 700)[%s PM]]',
          colors$ink_muted, format(e$effort_person_months, big.mark = ","))
}
typst(sprintf('
#grid(
  columns: (1fr, 1fr, 1fr, 1fr),
  gutter: 10pt,
  [#text(size: 8pt, fill: rgb("%s"))[COST] #linebreak() #text(size: 15pt, weight: 700, fill: rgb("%s"))[%s]],
  [#text(size: 8pt, fill: rgb("%s"))[SCHEDULE] #linebreak() #text(size: 15pt, weight: 700)[%s mo]],
  [#text(size: 8pt, fill: rgb("%s"))[TEAM] #linebreak() #text(size: 15pt, weight: 700)[%s people]],
  %s
)
#v(0.4cm)
',
  colors$ink_muted, colors$primary, esc(usd(e$cost_usd)),
  colors$ink_muted, e$final_schedule_months,
  colors$ink_muted, e$final_people,
  tco_cell
))
```

```{r warnings}
#| output: asis
if (!e$schedule_feasible) {
  cat(sprintf(
    "> **Note:** Delivering in %s months would require ~%s people, above the %d-person team cap. Consider extending the schedule.\n\n",
    e$final_schedule_months, e$final_people,
    as.integer(min(e$params$max_team_size, 8))
  ))
} else if (e$compression_penalty > 1.0) {
  cat(sprintf(
    "> **Note:** The schedule is compressed from a natural %s months; a +%d%% compression premium is included.\n\n",
    e$natural_schedule_months, round((e$compression_penalty - 1) * 100)
  ))
}
```

```{r waterfall}
#| fig-width: 6.6
#| fig-height: 3.1
gg_waterfall(e, colors)
```

```{r tornado}
#| fig-width: 6.6
#| fig-height: 3.1
gg_driver_tornado(e, colors)
```

```{=typst}
#pagebreak()
```

# Schedule & Ownership

```{r schedule-curve}
#| fig-width: 6.6
#| fig-height: 3.0
gg_schedule_curve(d$curve, e, colors)
```

```{r tco}
#| fig-width: 6.6
#| fig-height: 2.9
if (!is.null(e$maintenance)) gg_tco(e, colors)
```

```{r tco-note}
#| output: asis
if (!is.null(e$maintenance)) {
  m <- e$maintenance
  cat(sprintf(
    "Maintenance is projected at %.0f%% of build cost per year for %d years, escalating 5%%/yr. Total maintenance: %s nominal, %s in present value (discounted at %.0f%%/yr). **Total cost of ownership: %s.**\n\n",
    m$maintenance_rate * 100, m$maintenance_years,
    usd(m$total_maintenance_nominal), usd(m$total_maintenance_pv),
    m$discount_rate * 100, usd(m$tco)
  ))
}
```

## Assumptions

```{r assumptions}
#| output: asis
p <- e$params
bd <- d$billable
lines_label <- if (isTRUE(bd$has_docs)) {
  sprintf("%s billable (%s docs/config excluded)",
          format(bd$billable_code, big.mark = ","),
          format(bd$total_code - bd$billable_code, big.mark = ","))
} else {
  format(bd$billable_code, big.mark = ",")
}
rows <- rbind(
  c("Code lines", lines_label),
  c("Complexity", p$complexity),
  c("Team experience (1-5)", p$team_experience),
  c("Reuse factor", p$reuse_factor),
  c("Tool support", p$tool_support),
  c("Average annual wage (fully loaded)", usd(p$avg_wage)),
  c("Max team size", p$max_team_size),
  c("Max schedule (months)", p$max_schedule_months),
  c("Reliability (RELY)", p$rely),
  c("Product complexity (CPLX)", p$cplx),
  c("Reusability (RUSE)", p$ruse),
  c("Personnel continuity (PCON)", p$pcon),
  c("Application experience (APEX)", p$apex),
  c("Maintenance rate / years", sprintf("%.0f%% / %d", p$maintenance_rate * 100, p$maintenance_years)),
  c("Discount rate", sprintf("%.0f%%/yr", p$discount_rate * 100))
)
df <- as.data.frame(rows, stringsAsFactors = FALSE)
names(df) <- c("Parameter", "Value")
print(knitr::kable(df, format = "pipe"))
```

```{=typst}
#pagebreak()
```

# Methodology

This estimate is produced with a calibrated **COCOMO II.2000** model (Boehm et al., *Software Cost Estimation with COCOMO II*, 2000).

**Effort.** Person-months are estimated as

`Effort = 2.94 × KLOC^B × EM_total × 0.85`

where `KLOC` is thousands of productivity-weighted code lines, `B` is 1.02 / 1.10 / 1.18 for low / medium / high complexity, `EM_total` is the product of the effort multipliers listed in the assumptions table, and 0.85 is a documented calibration for modern high-level frameworks.

**Schedule.** The nominal schedule follows `TDEV = 3.67 × Effort^D` with `D = 0.28 + 0.2 × (B − 0.91)`. Team size is effort divided by schedule, capped at your team limit (and at 8 people, beyond which coordination costs outweigh added capacity).

**Schedule compression.** If the schedule cap is tighter than the natural schedule, effort is increased by a smooth premium `ratio^1.25` (anchored to COCOMO II's SCED driver), capped at 2.0×. Teams of 6 or more carry a 10% coordination overhead.

**Cost.** Cost is effort × monthly rate, where the monthly rate is the fully-loaded annual wage divided by 12. Non-code lines (documentation, configuration, markup) are excluded from billable size.

**Estimate range.** The ±30% band reflects typical COCOMO II estimation accuracy for uncalibrated projects. It is a heuristic accuracy band, not a statistical confidence interval. Actual cost depends on requirements stability, team specifics, and scope decisions outside this model.

**Maintenance.** Annual maintenance is a fixed share of build cost, escalating 5% per year; multi-year totals are reported both nominally and discounted to present value.

```{r footer}
#| output: asis
typst(sprintf('
#v(1fr)
#line(length: 100%%, stroke: 0.5pt + rgb("%s"))
#text(size: 8.5pt, fill: rgb("%s"))[This document was generated automatically from the parameters listed above. For questions about the estimate or to explore alternative scenarios, contact the report author.]
', colors$border, colors$ink_muted))
```
